---
title: "Assignment 4"
author: "Nond Prueksiri"
date: "April 8, 2019"
output: pdf_document
---

# Exercise 1 Data
```{r setup, include=FALSE}
rm(list = ls())
library(foreign)
library(ggplot2)
library(dplyr)
library(plm)
library(data.table)
library(dummies)
library(httr)
Koop.Tobias <- read.csv("https://raw.githubusercontent.com/ms486/Econ613/master/Assignments/A4/Koop-Tobias.csv")
panel <- pdata.frame(Koop.Tobias, index=c("PERSONID","TIMETRND"), drop.index=FALSE, row.names=TRUE)
panel_default <- pdata.frame(Koop.Tobias, index=c("PERSONID","TIMETRND"), drop.index=FALSE, row.names=TRUE)

sample <- sample(unique(panel$PERSONID), 5)
sample_df <- subset(panel, panel$PERSONID %in% sample)
as.matrix(sample_df$LOGWAGE)

```


# Exercise 2 Random Effects
```{r Random Effects, warning= FALSE}

rm(sample, sample_df)


# Step 1
OLS <- lm(LOGWAGE ~ EDUC + POTEXPER, data = panel)
summary(OLS)
var_alpha <- (coef(summary(OLS))[1,2])^2
var_resid <- var(OLS$residuals)
lamda <- 1 - sqrt(var_resid / (var_resid + 15*var_alpha))
lamda

# Step 2
panel <- data.table(panel)
setkey(panel,PERSONID)
panel <- panel[,mean_LOGWAGE:=mean(LOGWAGE),by=PERSONID]
panel <- panel[,mean_POTEXPER:=mean(POTEXPER),by=PERSONID]
panel <- panel[,mean_EDUC:=mean(EDUC),by=PERSONID]

panel <- panel[,re_LOGWAGE:= LOGWAGE - lamda*mean_LOGWAGE]
panel <- panel[,re_POTEXPER:= POTEXPER - lamda*mean_POTEXPER]
panel <- panel[,re_EDUC:= EDUC - lamda*mean_EDUC]

RE <- lm(re_LOGWAGE ~ re_POTEXPER + re_EDUC, data = panel)
summary(RE)

RE_check <- plm(LOGWAGE ~ POTEXPER + EDUC, data = panel_default, model = "random")
summary(RE_check)

```

# Exercise 3 Fixed Effects
```{r Fixed Effects, warning= FALSE}

#between estimator
Between <- lm(mean_LOGWAGE ~ mean_POTEXPER + mean_EDUC, data = panel)
summary(Between)

#within estimator
panel <- panel[,fe_LOGWAGE:= LOGWAGE - mean_LOGWAGE]
panel <- panel[,fe_POTEXPER:= POTEXPER - mean_POTEXPER]
panel <- panel[,fe_EDUC:= EDUC - mean_EDUC]
Within <- lm(fe_LOGWAGE ~ 0 + fe_POTEXPER + fe_EDUC, data = panel)
summary(Within)

#First time difference
panel <- panel[,d_LOGWAGE:=c(NA,diff(LOGWAGE)),by=PERSONID]
panel <- panel[,d_EDUC:=c(NA,diff(EDUC)), by=PERSONID]
panel <- panel[,d_POTEXPER:=c(NA,diff(POTEXPER)), by=PERSONID]
panel <- panel[,d_TIMETRND:=c(NA,diff(TIMETRND)),by=PERSONID]


FD <- lm(d_LOGWAGE ~ 0 + d_EDUC + d_POTEXPER, data = panel[d_TIMETRND == 1])
summary(FD)

```



# Exercise 4
```{r 4}

sample100 = sample(unique(panel$PERSONID), 100)
panel100 <- subset(panel_default, panel_default$PERSONID %in% sample100)
D <- dummy("PERSONID", data = panel100, sep="")
X <- panel100[,2]
X <- cbind(X, panel100[,4], D)
Y <- panel100[,3]

mle <- function(coeff) {
  beta <- coeff[1:102]
  sigma <- exp(coeff[103])
  p <- (Y - X%*%beta)/sigma
  ll <- sum(log(dnorm(p)/sigma))
  return(-ll)
}

coeff <- matrix(0, nrow = 103)

est <- nlm(mle,coeff)

panel_alpha <- as.matrix(est$estimate)[3:102,]
panel_alpha <- cbind(panel_alpha, sort(sample100))
FE_character <- unique(merge(panel_alpha, panel100, by.x ="V2", by.y="PERSONID")[,-(3:6)])
lm_Character <- lm(panel_alpha ~ ABILITY + MOTHERED +FATHERED + BRKNHOME + SIBLINGS, data=FE_character)


```

The standard error here may not be correct due to the fact that we random only 100 observation 
from the whole sample of more than 2,000 individual. The possible method to improve this is to
calculate them by "bootstrap" in order to increase the precision of standard errors given there are
more simulated datasets