---
title: "Assignment 4 Linear Panel Data"
author: "Nond Prueksiri"
date: "April 8, 2019"
output: pdf_document
---

# Exercise 1 Data
## Data Preparation
```{r setup, echo=FALSE}
rm(list = ls())
library(foreign)
library(ggplot2)
library(dplyr)
library(plm)
library(data.table)
library(dummies)
library(httr)
Koop.Tobias <- read.csv("https://raw.githubusercontent.com/ms486/Econ613/master/Assignments/A4/Koop-Tobias.csv")
panel <- pdata.frame(Koop.Tobias, index=c("PERSONID","TIMETRND"), drop.index=FALSE, row.names=TRUE)
panel_default <- pdata.frame(Koop.Tobias, index=c("PERSONID","TIMETRND"), drop.index=FALSE, row.names=TRUE)
```

## Panel dimension of wages for 5 randomly selected individuals
```{r 1.Sample Data}
sample <- sample(unique(panel$PERSONID), 5)
sample_df <- subset(panel, panel$PERSONID %in% sample)
as.matrix(sample_df$LOGWAGE)

```


# Exercise 2 Random Effects
To estimate random effect by hand, I do the two following steps;
* Step 1: estimate lamda ( 1 - ratio of standard error of residual over standard error of residual plus standard error
of intercept term times number of time periods) by pooled OLS

```{r 2A Random Effects, warning= FALSE}

rm(sample, sample_df)

# Step 1
OLS <- lm(LOGWAGE ~ EDUC + POTEXPER, data = panel)
var_alpha <- (coef(summary(OLS))[1,2])^2
var_resid <- var(OLS$residuals)
lamda <- 1 - sqrt(var_resid / (var_resid + 15*var_alpha))
lamda
```

* Step 2: use lamda estimated from step 1 to transform variable, that is; x_transform = x - lamda*mean_x

``` {r 2B Random Effects, warning= FALSE}
# Step 2
panel <- data.table(panel)
setkey(panel,PERSONID)
panel <- panel[,mean_LOGWAGE:=mean(LOGWAGE),by=PERSONID]
panel <- panel[,mean_POTEXPER:=mean(POTEXPER),by=PERSONID]
panel <- panel[,mean_EDUC:=mean(EDUC),by=PERSONID]

panel <- panel[,re_LOGWAGE:= LOGWAGE - lamda*mean_LOGWAGE]
panel <- panel[,re_POTEXPER:= POTEXPER - lamda*mean_POTEXPER]
panel <- panel[,re_EDUC:= EDUC - lamda*mean_EDUC]

RE <- lm(re_LOGWAGE ~ re_POTEXPER + re_EDUC, data = panel)
summary(RE)
```

* Check if the estimated random effects closed to the estimation provided by the package plm
```{r 2C Check}
RE_check <- plm(LOGWAGE ~ POTEXPER + EDUC, data = panel_default, model = "random")
summary(RE_check)

```

# Exercise 3 Fixed Effects

## Between Estimator
```{r 3A Fixed Effects, warning= FALSE}

between_panel <- aggregate(panel, list(panel$PERSONID), mean)
Between <- lm(LOGWAGE ~ EDUC + POTEXPER, data = between_panel)
summary(Between)
```

## within estimator
```{r 3B Fixed Effects, warning= FALSE}
panel <- panel[,fe_LOGWAGE:= LOGWAGE - mean_LOGWAGE]
panel <- panel[,fe_POTEXPER:= POTEXPER - mean_POTEXPER]
panel <- panel[,fe_EDUC:= EDUC - mean_EDUC]
Within <- lm(fe_LOGWAGE ~ 0 + fe_EDUC + fe_POTEXPER, data = panel)
summary(Within)
```

## First time difference
```{r 3C Fixed Effects, warning= FALSE} 

panel <- panel[,d_LOGWAGE:=c(NA,diff(LOGWAGE)),by=PERSONID]
panel <- panel[,d_EDUC:=c(NA,diff(EDUC)), by=PERSONID]
panel <- panel[,d_POTEXPER:=c(NA,diff(POTEXPER)), by=PERSONID]
panel <- panel[,d_TIMETRND:=c(NA,diff(TIMETRND)),by=PERSONID]
FD <- lm(d_LOGWAGE ~ 0 + d_EDUC + d_POTEXPER, data = panel[d_TIMETRND == 1])
summary(FD)
```
## Comparison between the three models
```{r 3D Comaprison}
FE <- coef(Between)
FE <- as.data.frame(rbind(FE, c(NA, coef(Within)) , c(NA,coef(FD))))  
row.names(FE) <- c("Between", "Within", "First time difference")
FE
```

# Exercise 4 Understanding Fixed Effects

* Write and optimize likelihood function to estimate "alpha" - individual fixed effects

```{r 4, warning= FALSE}

sample100 = sample(unique(panel$PERSONID), 100)
panel100 <- subset(panel_default, panel_default$PERSONID %in% sample100)
D <- dummy("PERSONID", data = panel100, sep="")
X <- panel100[,2]
X <- cbind(X, panel100[,4], D)
Y <- panel100[,3]

mle <- function(coeff) {
  beta <- coeff[1:102]
  sigma <- exp(coeff[103])
  p <- (Y - X%*%beta)/sigma
  ll <- sum(log(dnorm(p)/sigma))
  return(-ll)
}

coeff <- matrix(0, nrow = 103)

est <- nlm(mle,coeff)
```

* Regression of fixed effects
```{r 4B}
panel_alpha <- as.matrix(est$estimate)[3:102,]
panel_alpha <- cbind(panel_alpha, sort(sample100))
FE_character <- unique(merge(panel_alpha, panel100, by.x ="V2", by.y="PERSONID")[,-(3:6)])
lm_Character <- lm(panel_alpha ~ ABILITY + MOTHERED +FATHERED + BRKNHOME + SIBLINGS, data=FE_character)
summary(lm_Character)
```

* The standard errors
The standard error here may not be correct due to the fact that we random only 100 observation 
from the whole sample of more than 2,000 individual. The possible method to improve this is to
calculate them by "bootstrap" in order to increase the precision of standard errors given there are
more simulated datasets


